---
title: "Suicides Rates Analysis"
author: "Shu Wang"
date: '2022-08-08'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(rmarkdown)
library(lubridate)
library(countrycode)
library(rworldmap)
```


# Introduction
The suicide rate in lots of countries have been quite high for many years, including the developing and developed countries. In this study, we first investigated the distribution of suicides by age group. Studying the temporal changes in the number of suicides by age group can provide a more detailed understanding of suicide rates.

```{r}
#import the data
Data<-read.csv('master.csv')
Data<-Data%>%drop_na()
```


# Suicide Number of Different Age
```{r}
#per age
Data%>%
  group_by(age)%>%
  summarise(Suicide_Num=sum(suicides_no))%>%
  knitr::kable(caption="Suicide Number of Different Age")
```
From the above table, we can find that the number of suicides presents different characteristics according to different age groups. Those aged 35 to 54 had the highest number of suicides, while those aged 5-14 had the least.

# Data Visualization
## Temporal trends in the total number of suicides per year
```{r}
Data%>%
  group_by(year,suicides_no)%>%
  count()%>%
  mutate(num=n*suicides_no)%>%
  group_by(year)%>%
  summarise(Suicide_Num=sum(num))%>%
  ggplot(aes(x=year , y=Suicide_Num))+
  geom_line(color = "orange")+
  labs(title = "Suicide Number over the years")
```
From the graph above, we can see that the number of suicides increased rapidly from 1985 to 2000, and then stabilized and fluctuated after 2000.

## Suicide Number per age over the years
```{r}
Data$year <- as.factor(Data$year)

ggplot(Data, aes(year, suicides_no, fill=age )) +
  stat_summary(fun="mean", geom="bar") +
  facet_wrap(~age) +
  scale_fill_brewer(palette = "Set3")+
  labs(title = "Average Suicide Number per age over the years")
```
As can be seen from the above figure, the average number of suicides in each age group is relatively stable and shows significant differences. The youngest and oldest groups had the lowest total suicides, while the 35-54 and 55-75-year-olds had the highest average suicide rates.


# Conclusion
All in all, probably due to other factors such as life pressures brought about by social and economic development, suicide rates have been increasing worldwide since the 21st century. The age distribution trend of the average number of suicides remained relatively stable from 1985 to 2014.
